%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
  Among all human pursuits, the pursuit of wisdom is more \\
  perfect, more noble, more useful, and more full of joy. \medskip

  --- St. Thomas Aquinas \\
  (Summa Contra Gentiles, Book I, Chapter 2)}
  \\ \medskip
\end{flushright}



\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

Above all, I give thanks and praise to Almighty God,
the Most Holy Trinity, without whom nothing has meaning,
and without whose grace this journey would not have been possible.

To the Father, source of all life and providence, for
sustaining me each day. To Jesus Christ, my Savior and
Redeemer, for His mercy and love. To the Holy Spirit,
the Sanctifier, for shining light into the twofold darkness
I was born into: sin and ignorance.

\bigskip

To my father, Heberth de Paula, who has always been a model of
academic excellence and intellectual curiosity, and who has
endured great hardships and sacrifices to provide me
with the best education possible.

\bigskip

To my dearest, my love, and my sweetheart, Júlia de Paula, my
faithful companion, whose love have been the most precious
gift I've received in this passing world.

\bigskip

To my family, friends, work colleagues, and teachers, who
have supported me in countless ways, and who have always
believed in my potential.

\bigskip

To my former employer, Loggi, for taking a chance on me at 17
and investing in my growth—both as an engineer and a student—and
whose financial support enabled me to pursue and complete this
degree. I’m especially grateful to Davi de Castro Reis, who
incentivized me to pursue a degree in Computer Science and
whose support and decision enabled that investment on my education.

\bigskip

\centerline{\textcolor{gray}{\rule{6cm}{0.4pt}}}


\bigskip

\emph{May this work, in whatever good it contains, be for
the glory of God and the service of truth.}

\bigskip


\endgroup
